---
title: "SDS315HW8"
author: "Somya Krishna"
date: "2025-04-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE, 
                      warning = TRUE,
                      message = FALSE,
                      fig.align = "center", 
                      R.options = list(max.print=50))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load library
library(tidyverse)
library(mosaic)
```
https://github.com/somya-k535/SDS315HW8.git

# [Homework 8]{style="color:deeppink;"}

## [Problem 1: regression warm up]{style="color:hotpink;"}

#### [Part A]{style="color:pink;"}

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# A) What creatinine clearance rate should we expect for a 55-year-old? Explain briefly (one or two sentences + equations) how you determined this.

# read in data frame
creatinine <- read.csv("creatinine.csv")

# regression line
c_lm <- lm(creatclear ~ age, data = creatinine)
c_lm

# predict a 55-year old
ccr_55 <- 147.8129 - (0.6198*55)
ccr_55
```
**What creatinine clearance rate should we expect for a 55-year-old? Explain briefly (one or two sentences + equations) how you determined this.**

I would expect a creatinine clearance rate of 113.7239. I made a regression line using lm that stated:

creatine clearance rate = 147.8129 - (0.6198*age)

I then plugged in the age of 55 into the regression line formula.

#### [Part B]{style="color:pink;"}
**How does creatinine clearance rate change with age? (This should be a single number whose units are ml/minute per year.) Explain briefly (one or two sentences) how you determined this.**

Creatinine clearance goes down with age by 0.6198 ml/minute per year, on average. I determined this using a regression linear model.

#### [Part C]{style="color:pink;"}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# C) Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? Explain briefly (a few sentences + equations) how you determined this.
observed_40 <- 135
expected_40 <- 147.8129 - (0.6198*40)
print("Expected value for 40-year-old:")
expected_40
resid_40 <- observed_40 - expected_40
print("Residual for 40-year-old:")
resid_40
observed_60 <- 112
expected_60 <- 147.8129 - (0.6198*60)
print("Expected value for 60-year-old:")
expected_60
resid_60 <- observed_60 - expected_60
print("Residual for 60-year-old:")
resid_60

```
**Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? Explain briefly (a few sentences + equations) how you determined this.**

The observed rate for the 40-year-old is 135. The expected rate is 123.0209. This leaves us with a residual of 11.9791. The observed rate for the 60-year-old is 112. The expected rate is 110.6249. This leaves us with a residual of 1.3751. This means that the 40-year-old is healthier because their creatinine rate is higher for their age compared to the 60-year-old. I determined this by plugging in 40 and 60 into the regression line formula to get the expected values, and getting the residual by doing observed minus expected for each age.

## [Problem 2: Modeling disease growth]{style="color:hotpink;"}

#### [Part 1]{style="color:pink;"}
An estimated growth rate and doubling time for Italy, with 95% bootstrapped confidence intervals for
each.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# read in data frame
covid <- read.csv("covid.csv")

#Your task is to fit two exponential growth models, one for Italy and one for Spain, using days_since_first_death as the time variable. Use the results of your model to characterize the growth rate and doubling time of the daily death total in each country.

# Get only Italy
italy_data <- covid |>
  filter(country == "Italy")
# Bootstrap model
boot_italy_model <- do(10000)*lm(log(deaths) ~ days_since_first_death, data = resample(italy_data))
# Get growth rate for italy
italy_growth_rate <- confint(boot_italy_model)
italy_growth_rate <- round(italy_growth_rate$estimate[2], 3)
print("Italy Growth Rate:")
italy_growth_rate
# doubling time = log2/growth rate
italy_doubling_time = round(70/(italy_growth_rate*100), 1)
print("Italy Doubling Time:")
italy_doubling_time

# 95% bootstrapped confidence intervals for each.
# filter by x variable
#for growth rate
print("Confidence Interval for Italy's Growth Rate:")
confint(boot_italy_model, level = 0.95) |>
  filter(name=="days_since_first_death")

# ci for doubling time
print("Confidence Interval for Italy's Doubling Time:")
ci_upper_italy <- 70/(0.1588197*100)
print("Upper Bound")
ci_upper_italy
ci_lower_italy <- 70/(0.2077937*100)
print("Lower Bound")
ci_lower_italy
```
#### [Part 2]{style="color:pink;"}
An estimated growth rate and doubling time for Spain, with 95% bootstrapped confidence intervals for
each.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get only Spain
spain_data <- covid |>
  filter(country == "Spain")
# Bootstrap model
boot_spain_model <- do(10000)*lm(log(deaths) ~ days_since_first_death, data = resample(spain_data))
# Get growth rate for italy
spain_growth_rate <- confint(boot_spain_model)
spain_growth_rate <- round(spain_growth_rate$estimate[2], 3)
print("Spain Growth Rate:")
spain_growth_rate
# doubling time = log2/growth rate
spain_doubling_time = round(70/(spain_growth_rate*100), 1)
print("Spain Doubling Time:")
spain_doubling_time

# 95% bootstrapped confidence intervals for each.
print("Confidence Interval for Spain's Growth Rate:")
confint(boot_spain_model, level = 0.95) |>
  filter(name=="days_since_first_death")

# ci for doubling time
print("Confidence Interval for Spain's Doubling Time:")
ci_upper_spain <- 70/(0.2346624*100)
print("Upper Bound")
ci_upper_spain
ci_lower_spain <- 70/(0.3181369*100)
print("Lower Bound")
ci_lower_spain
```
#### [Part 3]{style="color:pink;"}
A line graph showing reported daily deaths over time (using days_since_first_death, rather than
calendar date, as the relevant time variable) in each country. Your line graph should have two lines,
one for each country, distinguished by their color.
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# line graph
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(title = "Daily Deaths Over Time for Italy vs. Spain",
       x = "Days Since First Death",
       y = "Daily Deaths",
       color = "Country")

```


## [Problem 3: price elasticity of demand]{style="color:hotpink;"}
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# read in data frame
milk <- read.csv("milk.csv")

boot_model <- do(10000)*lm(log(sales) ~ log(price), data=resample(milk))
print("Confidence Interval for elasticity of demand for milk 95% bootstrapped:")
confint(boot_model, level = 0.95) |>
  filter(name=="log.price.")

```
In light of the data, what is the estimated price elasticity of demand for milk? Give a 95% bootstrapped
confidence interval for this quantity. Briefly describe what you did to estimate the elasticity – no more than
a few sentences, together with your estimate + interval.

I made a regression model using the power law and took the log of sales and the log of price. I then bootstrapped it and for a 95% confidence level and filtered it to log.price to get the confidence interval of [-1.772865, -1.455241].
